---
title: "STAT 33B Homework 7"
author: "Gunnar Mayer (3034515354)"
output: pdf_document
---

This assignment is due __May 6, 2020__ by 11:59pm.

Edit this file, knit to PDF, and:

* Submit the Rmd file on bCourses.
* Submit the PDF file on Gradescope.

If you think you'll need help with submission, please ask in office hours
_before_ the assignment is due.

Answer all questions with complete sentences, and put code in code chunks. You
can make as many new code chunks as you like. Please do not delete the
exercises already in this notebook, because it may interfere with our grading
tools.


Return of the Bay Area Apartments Dataset
=========================================

The exercises in this assignment reuse the the Bay Area Apartments Data Set.

Please make sure to download the **new version** of the the data set from the
bCourse for this assignment.

You can find a full description of the data set in Lab 3.



Exercise 0
----------

Put all of your calls to `library()` here:
```{r, exercise0}
# Your code goes here.
library("dplyr")
```



Exercise 1
----------

Use dplyr to compute subsets and answer the following:

1. How many apartments are listed in Alameda?
2. How many of those apartments in Alameda have in-unit laundry?
3. What proportion of apartments between 800 and 1000 square feet (inclusive)
   have a rent price below 2200 dollars?

```{r, exercise1}
# Your code goes here.
ba_data = readRDS("apartments.rds")
alameda_data = filter(ba_data, county == "Alameda")

num_apart = summarise(alameda_data,n = n())
num_apart

num_laundry = summarise(filter(alameda_data, laundry == "in-unit"), n = n())
num_laundry

total_size = between(ba_data$sqft, 800, 1000)
price_filter = summarize(filter(ba_data, sizes, price <2200), n=n())
price_filter
total_size_sum = sum(total_size, na.rm = TRUE)
total_size_sum
ans = price_filter / total_size
ans

```

YOUR WRITTEN ANSWERS GO HERE:

1. There are 2321 apartments in Alameda county
2. There are 863 apartments with in-unit laundry in Alameda county
3.Less than 33% of the apartments in the Bay Area that are between 800 and 100 sqft are less than $2200. The proportion specifcally is 431 / 1312



Exercise 2
----------

Use dplyr's `count()` function to compute a data frame that shows the number of
bedrooms versus the number of bathrooms for all apartments in Oakland and San
Francisco.

_Hint: You solved this exercise with base R functions in Lab 3._

```{r, exercise2}
# Your code goes here.

filter(ba_data, county == "San Francisco" | county == "Oakland") %>% group_by(bedrooms, bathrooms) %>% count() 
```



Exercise 3
----------

Use dplyr to compute the maximum apartment price for each city, and sort your
results. What are the 3 cities with the highest maximum apartment price?

_Hint: See dplyr's online documentation for examples of how to use the
`group_by()`, `summarize()`, and `arrange()` functions._

```{r, exercise3}
# Your code goes here.
ba_data %>% group_by(city) %>% summarize(max_price = max(price)) %>% arrange(desc(max_price))

```

YOUR WRITTEN ANSWER GOES HERE:
The cities with the highest max rent are:

Berkeley ||	$16200			
San Francisco	|| $9500			
Oakland	|| $8900	


Exercise 4
----------

Compute the median apartment price (ignoring missing values) for each of the
different parking options. Use each of the three different approaches to the
split-apply pattern:

1. With `split()` and `sapply()`.
2. With `tapply()`.
3. With dplyr's `group_by()` and `summarize()`.

Confirm that the numerical results from each are the same. Describe other
differences between the results, such as the type and class of the returned
object.

```{r, exercise5}
# Your code goes here.

# 1
d1 = split(ba_data$price, f= ba_data$parking)
ans1 = sapply(d1, median, na.rm = TRUE)
ans1
typeof(ans1)
class(ans1)
# 2
ans2 = tapply(ba_data$price, ba_data$parking, median) 
ans2
typeof(ans2)
class(ans2)
# 3
ans3 = ba_data %>% filter(!is.na(parking), !is.na(price)) %>% group_by(parking) %>% summarise(median(price))
ans3
typeof(ans3)
class(ans3)
```

YOUR WRITTEN ANSWER GOES HERE:
The first 2 ways return doubles as answers, however using dplyr returns a list.
The first class using sapply returns a numeric class. The second method returns an array class. Dplyr returns a dataframe. Using tapply was definetly the simplist way to do it.
