---
title: "STAT 33B Lab 4"
author: "Gunnar Mayer (3034535154)"
output: pdf_document
---

Edit this file, knit to PDF, and:

* Submit the Rmd file on bCourses.
* Submit the PDF file on Gradescope.

If you think you'll need help with submission, please ask during the lab.

Answer all questions with complete sentences, and put code in code chunks. You
can make as many new code chunks as you like. Please do not delete the
exercises already in this notebook, because it may interfere with our grading
tools.

As you work, you may find it helpful to be able to run your code. You can run a
single line of code by pressing Ctrl + Enter. You can run an entire code chunk
by clicking on the green arrow in the upper right corner of the code chunk. 

Knit the document from time to time to make sure that your code runs without
errors from top to bottom in a fresh R environment. 


Monte Carlo Integration
=======================

In this lab, you'll practice writing code that uses loops, and learn how to
time or "benchmark" your code in order to compare the performance of different
implementations.

Monte Carlo integration is a statistical technique for computing the area (or
integral) of an arbitrary shape. The idea of the technique is to draw a square
around the shape and randomly, uniformly sample points inside the square. Then
the ratio of total points inside the shape to total points inside the square
(that is, all points) is the same as the ratio of the shape's area to the
square's area. This makes it possible to compute the area of the shape.

For example, suppose we want to use Monte Carlo integration to find the area of
the unit circle centered at the origin. We can enclose the circle with a square
that has opposite corners (-1, -1) and (1, 1). The area of this square is 4. If
we sample `n` points, then we can compute the area of the circle as:

`4 * (number of points in circle) / n`

Since the technique is random, it will produce a different result with each
run. However, the larger the total number of sampled points `n` is, the closer
the result will be to the true area of the shape.

In R, you can use the `runif()` function to generate random uniform samples.
For instance, to sample one value between -2 and 10, you would write `runif(1,
-2, 10)`. You can use `runif()` to sample points for Monte Carlo integration by
sampling the x and y coordinates separately.

Exercise 1
----------

Implement Monte Carlo integration for the unit circle example using `runif()`
and a for-loop. Specifically, each iteration of your loop should:

1. Sample an x coordinate.
2. Sample a y coordinate.
3. Test whether the sampled point falls inside the unit circle (use the
   equation for a circle).
4. Update a running total of the number of points that fell inside the unit
   circle.

After the loop, use the formula above to compute the area of the circle. Test
your code for 1 million points, which should be sufficient to compute pi to at
least 2 digits.

```{r}
# Your code goes here.
n = 1000000
inside_tot = 0
for(i in 1:n) {
  x = runif(1, -1, 1)
  y = runif(1, -1, 1)
  if(sqrt(x**2 + y**2) < 1) {
    inside_tot = inside_tot + 1
  }
}
pi = 4 * inside_tot / n
pi

```


Exercise 2
----------

In Exercise 1, you may have noticed that it takes seconds or even  minutes
(depending on your code and computer) to compute the result using 1 million
points. Vectorization is generally faster than for-loops, and all of the steps
in the Monte Carlo integration can be vectorized.

Implement Monte Carlo integration for the unit circle example using vectorized
operations (without a for-loop). Note that:

* The `runif()` function can sample more than one value at a time.
* Testing whether a point falls inside the unit circle is arithmetic. All of
  R's arithmetic operators are vectorized.
* The `sum()` function can compute the number of `TRUE` values in a vector.

Once again, test your code for 1 million points.

```{r}
n = 100000

x = runif(n, -1, 1)
y = runif(n, -1, 1)
inside_pt = sqrt(x**2 + y**2) < 1

pie = 4 * sum(inside_pt) / n
pie

```



Benchmarking Code
=================

When measuring the performance of a snippet of code, it's important to time the
code more than once. Averaging timings from multiple runs ensures that the
timings are not biased by other processes running on your computer.

The microbenchmark package provides a function `microbenchmark` that collects
multiple timings (with sub-millisecond accuracy) of code automatically. You can
install the package with the code:
```{r, eval = FALSE}
install.packages("microbenchmark")
```

Once the microbenchmark package is installed, you can time an expression by
passing it to the `microbenchmark` function:
```{r}
library(microbenchmark)

# How long does it take to sample 100 observations from standard normal
# distribution?
microbenchmark(rnorm(100))
```

You can also pass multiple expressions for the `microbenchmark` function to
time:
```{r}
# Which is faster, sampling 100 observations from standard normal or from
# standard uniform?
microbenchmark(rnorm(100), runif(100))
```

Finally, you can time multiple lines of code by surrounding them in curly
braces `{ }`. If you're timing more than one snippet of code, make sure to
surround all of them with curly braces `{ }`, since the curly braces themselves
add a small amount of extra time to the computation. For example:
```{r}
# Is it slower to sample 100 strings than it is to sample 100 integers?
microbenchmark(
  { # Case 1
    options = c("a", "b", "c")
    sample(options, 100, replace = TRUE)
  },
  { # Case 2
    options = 1:3
    sample(options, 100, replace = TRUE)
  }
)
```

Read the documentation in `?microbenchmark` for more details about using the
function.


Exercise 3
----------

Use the `microbenchmark` function to benchmark your code from Exercise 1 and
Exercise 2. You can reduce the number of sampled points to 10,000 in the
benchmarked code.

Discuss your result in 1-3 sentences. On your computer, how much slower is the
for-loop implementation compared to the vectorized implementation?

Make sure your benchmarks are displayed in the notebook you submit.

```{r}
# Your code goes here.
benchmark = microbenchmark(
  {
n = 10000
inside_tot = 0
for(i in 1:n) {
  x = runif(1, -1, 1)
  y = runif(1, -1, 1)
  if(sqrt(x**2 + y**2) < 1) {
    inside_tot = inside_tot + 1
  }
}
pi = 4 * inside_tot / n
pi
},
{
n = 10000
x = runif(n, -1, 1)
y = runif(n, -1, 1)
inside_pt = sqrt(x**2 + y**2) < 1

pie = 4 * sum(inside_pt) / n
pie
})

benchmark
```

YOUR WRITTEN ANSWER GOES HERE:

On my computer the difference between the mean of the for loop and vectorized operations is 42856.06 microseconds. Vector operations are substantially faster than looping. On average the for loop is 30x slower than the vector implementation, 44315.812 / 1459.752 ~= 30.
