---
title: "STAT 33B Lab 3"
author: "Gunnar Mayer (3034535154)"
output: pdf_document
---

Edit this file, knit to PDF, and:

* Submit the Rmd file on bCourses.
* Submit the PDF file on Gradescope.

If you think you'll need help with submission, please ask during the lab.

Answer all questions with complete sentences, and put code in code chunks. You
can make as many new code chunks as you like. Please do not delete the
exercises already in this notebook, because it may interfere with our grading
tools.

As you work, you may find it helpful to be able to run your code. You can run a
single line of code by pressing Ctrl + Enter. You can run an entire code chunk
by clicking on the green arrow in the upper right corner of the code chunk. 

Knit the document from time to time to make sure that your code runs without
errors from top to bottom in a fresh R environment. 

The code below controls the number of significant digits shown for the return
values in your knitted document.

```{r}
options(digits = 3)
```

Datasaurus Dozen Data Set
=========================

Exercise 1 uses the Datasaurus Dozen Data Set from homework 2.


Exercise 1
----------

A "faceted" plot is one that shows several plots side-by-side, to aid
comparison between them. Each subplot is called a "facet".

You can create a faceted plot with `ggplot2` by using the facet layer. For
instance, the `facet_wrap()` function creates a line of facets based on a
single categorical variable. The facet layer should be added to a plot _after_
the geometry layers.

1.  Read the documentation for `facet_wrap()`, then create a faceted scatter
    plot that shows each dataset from the Datasaurus Dozen in a separate facet.
    Use `geom_smooth` with `method = "lm"` to add a linear regression line to
    each facet.

    _Hint: Unlike other `ggplot2` functions, the facet functions do not accept
    unquoted variable names. See the documentation for details._

2. Is there any pattern to the regression lines across the different data sets?

```{r}
# Your code goes here.
library(ggplot2)
dsaur = read.delim(file = "DatasaurusDozen.tsv")
ggplot(data = dsaur, aes(x = x, y = y)) +
    geom_point() + 
    geom_smooth(method = "lm") +
    facet_wrap(vars(dsaur$dataset))    
```

YOUR WRITTEN ANSWER GOES HERE: Is there any pattern to the regression lines across the different data sets?

Yes there is a pattern to the regression lines across the different data sets. All of the regression lines have very similar slopes. This means that the average of all the datasets inside dsaur will be relatively similar.

Bay Area Apartments Data Set
============================

For the remainder of this lab, you'll continue to analyze the Bay Area
Apartments Data Set introduced in lab 2.

Recall that the Bay Area Apartments Data Set is a collection of 5852
advertisements for apartments for rent in the San Francisco Bay Area. The data
set was collected from Craigslist on Feb 13, 2020.


Exercise 2
----------

Histograms, density plots, and box plots are all good ways to display the
distribution of a 1-dimensional numerical feature. You can use side-by-side box
plots to compare a distribution across the levels of a categorical feature, 

1.  Use `ggplot2` to create side-by-side box plots that show the distribution
    of apartment prices broken down by number of bathrooms. Add descriptive
    labels and a title to your plot.

    _Hint: The `ggplot2` package treats discrete data and categorical data
    differently. You can use the `factor()` function to convert a vector or
    column into categorical data._

2.  Based on your plot, does the number of bathrooms affect the price? If so,
    how?

```{r}
ba_data = readRDS("cl_apartments.rds")
ggplot(data = ba_data) +
    geom_boxplot(aes(x = bathrooms, y = price)) +
    facet_wrap(vars(ba_data$bathrooms)) +
    labs(title = "Price vs Bathroom Number")+
    labs(x = "Bathrooms") + labs(y = "Price")

```

YOUR WRITTEN ANSWER GOES HERE: Based on your plot, does the number of bathrooms affect the price? If so,
    how?
    
Based on the data having 3 bathrooms drives up the price of the unit. For the most part it doesn't matter too much how many bathrooms you have. As the number of bathrooms increased past 4.5 the price actually went down. I found this surprising however I confirmed this is the case by taking the sample of all apartments that had 6 bathrooms. Surprisingly there was only one, and that one was listed for only $800 in Berkeley.

Exercise 3
----------

Bar plots are a good way to display the distribution (that is, the frequencies)
of a 1-dimensional categorical feature. Groups of bars can be used to compare
against a second categorical feature.

1. Get the subset of advertisements for apartments with 0 to 3 bedrooms.

2. For the subset, create a bar plot that shows the frequencies of the different laundry
categories, 
    a. grouped by the number of bedrooms (on the x-axis). 
    b. Position the bars so that they are side-by-side rather than stacked. 
    c. Add descriptive labels and a title to your plot.

_Hint: You will need to set the `position` parameter for the geometry layer._

```{r}
bath_comp <- subset(ba_data, ba_data$bathrooms <= 3)
ggplot(data = bath_comp, ) +
    geom_bar(aes((laundry), group = 1), position = position_dodge2()) +
    labs(title = "Frequencies of the Different Laundry", x = "Number of Bathrooms", y = "Laundry Count")

```

Exercise 4
----------

The `table()` function computes the frequency of each unique value in a vector.
If you want to inspect the distribution of categories in a categorical feature,
the `table()` function is a quick alternative to a bar plot. 

Use the `table()` function to compute a two-way table that shows the number of
bedrooms versus the number of bathrooms for all of the advertisements in
Oakland and San Francisco. Use the `dnn` parameter to label the appropriate
dimensions of the table as "Bedrooms" and "Bathrooms".

Which combination of number of bedrooms and number of bathrooms is most common
among apartment advertisements for these two cities?


```{r}
sf_and_oak = subset(ba_data, ba_data$city == "San Francisco"| ba_data$city == "Oakland")
with(sf_and_oak, table(bedrooms, bathrooms), useNA = FALSE, dnn = "bedrooms",  "bathrooms")

```

YOUR WRITTEN ANSWER GOES HERE: Which combination of number of bedrooms and number of bathrooms is most common among apartment advertisements for these two cities?

Based off of the table I found that the most common combination of bathroom and bedroom is having one of each. 


Exercise 5
----------

Given a vector, the `order()` function returns the vector of positions that
sort the vector if used with the `[` operator. For example, suppose we want to
order this vector from smallest to largest:
```{r}
x = c(4, -6, 21, -2)
```

We can get the ordered vector by calling `order()` and then subsetting:
```{r}
indices = order(x)
indices

x[indices]
```

The `order()` function generalizes to data frames. If you get the ordering for
a column, you can then use the ordering to subset the rows and reorder them.

Use `order()` to find the 5 least expensive apartment advertisements in
Berkeley. Print their price, number of bedrooms, and title. Are all of them ads
for apartments?

```{r}
berk <- subset(ba_data, ba_data$city == 'Berkeley')
berk <- na.omit(berk)
cheap_berk <- order(berk$price)
cheap_berk <- cheap_berk[0:5]
r <- c(0,1,2,3,4)
for (i in r){
    print(berk$price[cheap_berk[i]])
    print(berk$bedrooms[cheap_berk[i]])
    print(berk$title[cheap_berk[i]])
}
```

YOUR WRITTEN ANSWER GOES HERE: Are all of them ads for apartments?

Yes, they appear to all be ads for apartments. I'm not sure why the price is NA even after I omitted NAs.
