---
title: "Stat 33B - Lecture 7"
date: March 4, 2020
output: pdf_document
---

Announcements
=============

Due to the coronavirus, I'm considering alternatives to in-person
lectures/labs and the attendance grade.

Homework 4 will be posted tomorrow.


Review
======

Grocery data set on bCourse:
```{r}
stores = readRDS("data/grocery/stores.rds")
items = readRDS("data/grocery/items.rds")
inv = readRDS("data/grocery/inventory.rds")
```

A merge or "join" combines data from two separate data frames, based
on some identifying values they have in common:
```{r}
merge(items, inv)
```


If-statements
=============

If-statements run some code only if a condition is true.

Three ways to write if-statements in R.

1. Ordinary if-statements

```{r}
# You have to have the brackets around if and else in exactly the same way
x = 5
if(x<5) { message("x is less than 5!") 
} else { message("this is the else" )}
# White space doesn't matter in R, so one liner ifs can be on one line or however we want

```

Multiple if-statements can be chained together:

```{r}
if (){  }

else if () {  }

else {  }
```

Use `{`, the curly brace, to write multiple line if-statements:
```{r}

```

If-statements return the value of their last expression:

```{r}
# Conditional statements return whatever the last thing in the block is

x = 3
y = if ( x < 4) 42 else 8
y

# This sets y = 42, or 8 if x >= 4
```


2. The `ifelse()` function is a vectorized if-statement.
  
```{r}
# This type of if returns an error
x = c(1,2,3)
if(x<2) {
  message("x<2")
}

# The ifelse() function goes through the entire vector
# This replaces 10 for values >= 2, can be used to change values of a vector through reassignment
ifelse(x < 2, x, 10)
x
```

Avoid `ifelse()` unless you need vectorization. Slower than ordinary
if-statement.


3. The `switch()` function

Matches an integer or string to one of several possible branches.

```{r}
#how to write swtich statements
switch(#thing it is testing, #case1, #case2, ...,
)
# a real example, R matches the string in x to the cases in the switch statement
x = "hello"

switch(x, 
       "hello" = message("You said hello")
       "hye" = message("You said bye")
       "33b"  = message("33b")
       
# you can have a catchall in the switch statement by not checking whether something is true, ie below
       message("This is a catch all")
       )
# Only strings and integers work in this, integers work based off of position
x = 2
switch(x,
  message("x is 1"),
  message("x is 2")
)

# if you use a string as if it were an integer it will throw an error unless there is only one case
x = "hi"
switch(x,
  message("x is 1"),
  message("x is 2")
)

#Error: duplicate 'switch' defaults: 'message("x...' and 'message("x...'
#it things everything is a default
```

You can also do this with a list:

```{r}
# if you put something in your list that is very time consuming it will always run
# if you put a time consuming function in your switch CASES then it won't always run
mycases  = list(1, 30)
mycases[[x]]
```

What's the difference?


Loops
=====

Loops run some code repeatedly.

Four ways to write loops in R:

1. Vectorization (fast!)
2. Apply functions: `lapply()`, `sapply()`, etc
3. `for`, `while`, and `repeat` loops
4. Recursion (slow)

Vectorization
-------------

A "vectorized" function is one that automatically applies an
operation to each element of a vector.

We saw this earlier with arithmetic:
```{r}
# This is a loop and the fastes way to loop
sin(c(1, 2, 3))
```

Apply Functions
---------------

Similar to "map" functions in other languages.

Applies a function (FUN) to each element of a vector or list (X):
```
lapply(X, FUN, ...)
```


`lapply()` always returns the result as a list:
```{r}
# applies a function to each element in a list
# this is good for functions that aren't vectorized, but you want to apply to the entire list/vector

#lapply function lapply(data, function)

#input does NOT have to be a list, but it always returns a list
mydata = list(1:3, seq(10, 20, 0.5), rnorm(10))
mydata

lapply(mydata, mean)

# how to remove NA, adding this function is fixed for all iterations. Adding anything extra is applied to everything

lapply(mydata, mean, na.rm = TRUE)
```
"l" for **list**.


Most apply functions will accept additional arguments to the applied
function. These arguments are held constant across all elements of
the data:
```{r}
dog = readRDS("/Users/g-mayer/adv_r/lab/lab_1/dogs_full.rds")

# lapply keeps the names of the input
lapply(dog, class)
```


`sapply()` simplifies the result to a vector, when possible:
```{r}
# it might return a list or it might return a vector
sapply(list, function)
```
"s" for **simplified** result.


`vapply()` simplifies the result to a vector of a specific type:
```{r}
# always returns a vector 
# vapply(data, function, an example of what will be returned )
dog = readRDS("/Users/g-mayer/adv_r/lab/lab_1/dogs_full.rds")

vapply(dog, class, '')
```
"v" is for **vector** result.



`mapply()` applies a function to each element of multiple vectors or
lists:
```{r}
# looping over multiple vectors at once

x = c(1,2,3)
y = 2:4

# the best way to do it, using vectorization
x + y
############

# mapply(function, data1, data2)
mapply(`+`, x, y)
```
"m" for **multiple** vectors.


### Flashback to Data Analysis

The `split()` function splits a vector or data frame into groups
based on some other vector.


```{r}

```

The `split()` function is especially useful when combined with
`lapply()` or `sapply`().

```{r}

```
This is an R idiom!


The `tapply()` function is equivalent to the `split()` and `sapply()`
idiom:
```{r}

```
"t" for **table**, because `tapply()` is a generalization of the
frequency-counting function `table()`.


The `split()` and `tapply()` functions are important for analyzing
data in data frames.


### Other Apply Functions

* `apply()` applies a function to rows, columns, or elements of a
  matrix, and returns a matrix.
* `eapply()` applies a function to the elements of an environment.
  We'll learn about environments in a later lecture.
* `rapply()` applies a function to all elements of a recursive list
  (a list of lists).

See this StackOverflow Post for a good summary:

    https://stackoverflow.com/a/7141669



`for`, `while`, and `repeat`
----------------------

A for-loop runs a block of code once for each element of a vector or
list:
```{r}

```
The same idea as for-loops in other languages.


A while-loop runs a block of code repeatedly as long as some
condition is true:
```{r}

```
Again, the same idea as in other languages.


Use `break` to exit a loop early:
```{r}

```

Use `next` to skip to the next iteration early:
```{r}

```


A repeat-loop runs a block of code repeatedly:
```{r}

```

If you know how many results you expect to get from the loop,
allocate storage before the loop begins:
```{r}

```

If you need indices, using `1:length(x)` can cause bugs:
```{r}

```

Use `seq_along(x)` instead:
```{r}

```


Recursion
---------

A recursive function is one that calls itself:
```{r}

```

R's built-in `Recall()` function will re-call the enclosing function:
```{r}

```

Without `Recall()`, renaming a recursive function can break the
function:
```{r}

```


How to Write Loops
------------------

When thinking about writing a loop, try (in order):

1. vectorization
2. apply functions
    * Try an apply function if iterations are independent.
3. for/while-loops
    * Try a for-loop if some iterations depend on others.
    * Try a while-loop if the number of iterations is unknown.
4. recursion
    * Convenient for naturally recursive problems (like Fibonacci),
      but often there are faster solutions.

Before you write the loop, try writing the code for just 1 iteration.

Make sure that code works; it's easy to test code for 1 iteration.

When you have 1 iteration working, then try using the code in a loop
(you will have to make some small changes).

If your loop doesn't work, try to figure out which iteration is
causing the problem. One way to do this is to use message() to print
out information.

Then try to write the code for the broken iteration, get that
iteration working, and repeat this whole process.


Functions
=========

You can write your own function with the `function` command.

Just like any other object in R, you can assign functions to
variables:
```{r}

```

To write a function with more than one line of code, use curly braces
`{ }`.

The last expression in the function is the output:
```{r}

```

Use `return()` to return a result earlier:
```{r}

```
It's idiomatic to only use `return()` when strictly necessary.


How to Write Functions
----------------------

Functions are the building blocks for solving problems.

Take a divide-and-conquer approach, breaking large problems into
smaller steps.

Use a short function for each step. Then it's easier to:

* Test that each step works correctly.
* Modify, reuse, or repurpose a step.

Before you write a function, it's useful to go through several steps:

1. Write down what you want to do, in detail. It can also help to
   draw a picture of what needs to happen.

2. Check whether there's already a built-in function. Google it!

3. Write the code to handle a simple case first. For data science
   problems, use a small dataset at this step.
